---
title: "Improving Visualizations"
author: "Martin Sloley"
output:
    bookdown::html_document2:
        theme: united
        toc: true
---

```{r setup, echo = F, message = F, warning = F}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
library(tidyverse)
```

Lately I have been working a lot on improving my technical skills, but I also need to be better about some of my soft skills, and in particular, my ability to communicate visually with data. In an effort to improve this ability, I am taking data from older projects and improving old visualizations, as well as adding new ones.

# Traffic in Seattle

We may as well start with the inspiration for this project, the [Scientific American "Baby Spike" visualization](https://www.visualcinnamon.com/2017/10/creating-baby-births-visual/). While reading Bremer's linked blog post, I realized that I had done a project that also contained cyclical time series information that may benefit from a visualization like this. We see the result of my attempt below.

```{r, message = F, fig.align = "center", fig.width = 10, fig.height = 10}
data = read_csv("data/seattle_collision_data_2005_2019.csv")
to_mins = function(mildectime) {
  mins = as.integer(60 * mildectime)
  return(mins)
}
by_mins2 = data %>%
        mutate(TIME = to_mins(TIME)) %>%
        group_by(DATE, TIME) %>%
        summarize(count = n()) %>%
        group_by(TIME) %>%
        summarize(mean_cpm = mean(count))

by_mins = data %>%
        mutate(TIME = to_mins(TIME)) %>%
        group_by(TIME) %>%
        summarize(count = n()) %>%
        mutate(total_mean = mean(count),
               rolling_mean = DescTools::MoveAvg(count, 5, "center")) %>%
        mutate(norm_dist_from_mean = (rolling_mean - total_mean) / sd(count)) %>%
        arrange(TIME)

mod = lm(norm_dist_from_mean ~ poly(TIME, 20), by_mins)
by_mins$fit_vals = predict(mod, by_mins)

p1 = ggplot(by_mins) +
        geom_ribbon(aes(x = TIME, y = ifelse(fit_vals >= 0, fit_vals, 0),
                        ymax = ifelse(fit_vals >= 0, fit_vals, 0), ymin = 0),
                    alpha = 0.6, fill = "#EF5945") +
        geom_ribbon(aes(x = TIME, y = ifelse(fit_vals < 0, fit_vals, 0),
                        ymin = ifelse(fit_vals < 0, fit_vals, 0), ymax = 0),
                    alpha = 0.8, fill = "#18C3EA") +
        geom_line(aes(x = TIME, y = rep(0, nrow(by_mins))),
                  color = "#EFE745", size = 1.5) +
        geom_line(aes(x = TIME, y = fit_vals), size = 0.2) +
        geom_point(aes(x = TIME, y = norm_dist_from_mean),
                   alpha = 0.3, shape = 1, size = 0.8) +
        labs(x = "Vehicle Collisions\nper Minute in the Day") +
        scale_y_continuous(limits = c(-2, 2.5)) +
        scale_x_continuous(limits = c(-120, 1560)) +
        geom_segment(aes(x = 10, y = 0, yend = 1, xend = 10), size = 0.3) +
        annotate("text", x = 10, y = 1.2, label = "Average Collision Rate") +
        geom_segment(aes(x = 1020, y = predict(mod, data.frame(TIME = 1020)),
                         yend = 2.2, xend = 1020),
                     linetype = "dashed") +
        geom_segment(aes(x = 540, y = predict(mod, data.frame(TIME = 540)),
                         yend = 2.2, xend = 540),
                     linetype = "dashed") +
        geom_segment(aes(x = 4 * 60, y = predict(mod, data.frame(TIME = 4 * 60)),
                         yend = 2.2, xend = 4 * 60),
                     linetype = "dashed") +
        annotate("text", x = 540, y = 2.5, label = "9 am") +
        annotate("text", x = 1020, y = 2.5, label = "5 pm") +
        annotate("text", x = 4 * 60, y = 2.5, label = "4 am") +
        coord_polar() +
        theme(axis.line.y = element_blank(),
              panel.grid.major.x = element_line(size = 1),
              panel.grid.major.y = element_blank(),
              panel.grid.minor = element_blank(),
              panel.border = element_blank(),
              panel.background = element_blank(),
              axis.text = element_blank(),
              axis.ticks.y = element_blank(),
              axis.title.y = element_blank(),
              axis.ticks.x = element_blank(),
              axis.title.x = element_text(size = 12, vjust = 120, face = "bold"))
p1
```
