---
title: "Improving Visualizations"
author: "Martin Sloley"
output:
    bookdown::html_document2:
        theme: united
        toc: true
---

```{r setup, echo = F, message = F, warning = F}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
library(tidyverse)
```

Lately I have been working a lot on improving my technical skills, but I also need to be better about some of my soft skills, and in particular, my ability to communicate visually with data. In an effort to improve this ability, I am taking data from older projects and improving old visualizations, as well as adding new ones.

# Traffic in Seattle

We may as well start with the inspiration for this project, the [Scientific American "Baby Spike" visualization](https://www.visualcinnamon.com/2017/10/creating-baby-births-visual/). While reading Bremer's linked blog post, I realized that I had done a project that also contained cyclical time series information that may benefit from a visualization like this. We see the result of my attempt below.

```{r, message = F}
data = read_csv("data/seattle_collision_data_2005_2019.csv")
to_mins = function(mildectime) {
  mins = as.integer(60 * mildectime)
  return(mins)
}

by_mins = data %>%
        mutate(TIME = to_mins(TIME)) %>%
        group_by(TIME) %>%
        summarize(count = n()) %>%
        mutate(total_mean = mean(count))

ggplot(by_mins) +
        geom_point(aes(x = TIME, y = count)) +
        geom_smooth(aes(x = TIME, y = count), method = loess) +
        geom_line(aes(x = TIME, y = total_mean)) +
        coord_polar()

mean(by_mins$count)
```
